---
title: "galah<br><h2>Workshop</h2><br>"
author: "Atlas of Living Australia"
date: 2022-11-25
title-block-banner: "#f5b2d0"
title-block-banner-color: "#222322"
title-block-style: plain
format: 
  html:
    css: style.css
    highlight-style: tango
---

## An introduction to galah

Install

```{r}
#| eval: false
#| message: false
#| warning: false
#| error: false
install.packages("galah")
install.packages("dplyr")
```

Load packages

```{r}
#| message: false
#| warning: false
library(galah)
library(dplyr)
```



### Need more info?

```{r}
#| eval: false
?galah
```



# Atlases

See what atlases are available:

```{r}
show_all(atlases)
```


Change atlas

```{r}
galah_config(atlas = "Austria")
```



# Fields

See all possible fields for filtering queries

```{r}
show_all(fields)
```

We recommend using the `View()` function, or `tibble::view()` to open see this information in a new tab within R Studio

```{r}
#| eval: false
show_all(fields) |> View()
show_all(fields) |> tibble::view()
```

Search for fields

```{r}
search_all(fields, "australian states") # not case-sensitive
```

See values within a field

```{r}
#| eval: false
search_all(fields, "cl22") |> 
  show_values()
```

Search for values within a field

```{r}
#| eval: false
search_all(fields, "cl22") |> 
  search_values("tas")
```



### Narrowing results ###

search_taxa("reptilia")

#add a dataframe

search_taxa(data.frame(genus="Eolophus", kingdom="Aves"))

#find numbers of records for all reptilia
galah_call() %>% 
  galah_identify("reptilia") %>% 
  atlas_counts()

#create an object
taxa<-search_taxa(data.frame(genus="Eolophus",kingdom="Aves"))

#use object in call
galah_call() %>% 
  galah_identify(taxa) %>% 
  atlas_counts()

###galah filter

#total records on ala since 2000
galah_call() %>% 
  galah_filter(year>2000) %>% 
  atlas_counts()

#all records since 2000 form iNaturalist australia
galah_call() %>% 
  galah_filter(
    year>=2000,
    dataResourceName=="iNaturalist Australia"
  ) %>% 
  atlas_counts()

#search fields values

search_field_values("dataResourceName",limit=1000) %>% View()

### profiles

#full list of profiles
show_all_profiles() %>% View()

#whats happening inside a profile
search_profile_attributes("ALA") %>% View()

#records form 2000 that are spatially valid
galah_call() %>% 
  galah_filter(year>=2000,
               profile = "ALA") %>% 
  atlas_counts()

###galah group by

#records for galah since 2015-2021 group by year and basis of record

```{r}
galah_config(email = "your-email-here") # ALA-registered email
```

```{r}
#| include: false
galah_config(email = "martin.westgate@csiro.au", verbose = FALSE)
```

galah_occurrence_data<-galah_call() %>% 
  galah_identify("Eolophus roseicapilla") %>% 
  galah_filter(year>2015 & year<=2021) %>% 
  galah_group_by(year, basisOfRecord) %>% 
  atlas_occurrences()
